---
title: "Comcast_Project"
author: "Michael Amanor Odonkor"
date: "12/24/2021"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---



# Comcast Telecom Consumer Complaints (Project 2)
### Load Libraries
```{r loading packages , results='hide',message=FALSE,warning=FALSE}
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(ggfortify)
library(scales)
library(knitr)

```



# Import data into R environment.
```{r}
raw_data <- read.csv("/Users/Manor/Desktop/Comcast Telecom Complaints data.csv", stringsAsFactors = FALSE)
```

### Cleaning the date format
```{r}
raw_data$Date <-dmy(raw_data$Date)
raw_data$Date <- as.Date(raw_data$Date)
```

### Working with sample data (removing columns we don't need)
```{r}
sample_data <-raw_data %>% 
  select(-c("Ticket..","Zip.code","Filing.on.Behalf.of.Someone","Time","City"))

```



# Provide the trend chart for the number of complaints at monthly and daily granularity levels.
### Daily and Monthly complaints
```{r}
daily_complaints <- sample_data %>%
  mutate(Daily_Complaints = strftime(sample_data$Date,"%d")) %>% 
  count(Daily_Complaints) %>% 
  rename(daily_Count = "n")

monthly_complaints <- sample_data %>% 
  mutate(Monthly_Complaints = strftime(sample_data$Date,"%m")) %>% 
  count(Monthly_Complaints) %>% 
  rename(monthly_Count = "n")
```

### Total daily complaint graph
```{r}
(g1 <- ggplot(data = daily_complaints,aes(x = Daily_Complaints, y = daily_Count , group = 1, color = "orange")) + geom_line() + geom_point() + labs (x = "Days", y = "Sum of Complaints", title = "Figure 1: A graph showing the daily complaints by customers", fill = NULL  ) + geom_path()+ theme_classic() + theme(legend.position="none"))

```

### The top 3 days with the highest number of complaints in a month are 6th, 24th and 23rd





### Total monthly complaint graph
```{r}
(g2 <- ggplot(data = monthly_complaints,aes(x = Monthly_Complaints, y = monthly_Count , group = 1, color = "orange")) + geom_line() + geom_point() + labs (x = "Months", y = "Sum of Complaints", title = "Figure 2 : A graph showing the monthly complaints by customers", fill = NULL  ) + geom_path()+ theme_classic() + theme(legend.position="none"))

```

### This chart shows that the Sum of complaints are highest in the month of June.





# Provide a table with the frequency of complaint types.
### Creating a new column with NA values
```{r}
sample_data <- sample_data %>%  
  add_column(Complaint_Type = NA)

```

### Classifying customer complaints into  5  groups using certain keys 
```{r}
internet <- contains(raw_data$Customer.Complaint ,match = c("internet service","gb", "bundle","cap","mb","speed","internet"),ignore.case = TRUE)
billing <- contains(raw_data$Customer.Complaint ,match = c("billing","charged", "charges","billed"),ignore.case = TRUE)
service <- contains(raw_data$Customer.Complaint ,match = "service",ignore.case = TRUE)
email <- contains(raw_data$Customer.Complaint ,match = "email",ignore.case = TRUE)

```

### Adding new classifications into the dataframe column as values 
```{r}
sample_data$Complaint_Type[internet] <- "Internet"
sample_data$Complaint_Type[service] <- "Service" 
sample_data$Complaint_Type[billing] <- "Billing"
sample_data$Complaint_Type[email] <- "Email"
sample_data$Complaint_Type[-c(internet,service,billing, email)] <- "Others"

```

### Frequency of Customer Complaints 
```{r}
Frequency_data <- data.frame(table(sample_data$Complaint_Type))
colnames(Frequency_data) <- c("Complaint_Type", "Frequency")
Sorted_frequency <- Frequency_data[order(-Frequency_data$Frequency),]
knitr::kable(Sorted_frequency, "pipe", align=c("c", "c"))
```

# Which complaint types are maximum?
### From the table with the frequency of complaint types above, we can see that the maximum complaint type is Others (779), followed by Internet, Billing, Service, and Email.





## Create a new categorical variable with value as Open and Closed. Open & Pending is to be categorized as Open and Closed & Solved is to be categorized as Closed.
### Create new column with NA values
```{r}

sample_data <- sample_data %>%  
  add_column(New_Status = NA)

```

### Categorize the data based on the status of complaint 
```{r}
sample_data$New_Status[sample_data$Status == "Open"|sample_data$Status == "Pending"] = "Open"
sample_data$New_Status[sample_data$Status == "Closed"|sample_data$Status == "Solved"] ="Closed"

```




# Provide state wise status of complaints in a stacked bar chart
### Finding the count of complaints based on the state and status
```{r}
state_wise_status <- sample_data %>% 
  group_by(State, New_Status) %>% 
  summarise(Count = n()) 
```

### Plotting Chart (stacked bar chart)
```{r}

(g3 <- ggplot(data = state_wise_status, aes(x = State, y = Count, fill = New_Status)) + 
    geom_bar(stat = "identity") +  coord_flip() + scale_fill_brewer(palette = "Dark2"))

(g3 + labs(title = "Figure 3  Shows the State Wise Status of Complaints", 
          y = "Number of Complaints", x = "States", fill = "Status"))
```


# Which state has the maximum complaints? 
### State with the most complaints in descending order
```{r}
Maximum_complaints <- data.frame (sample_data %>% 
                                    group_by(State) %>% 
                                    summarise(Count = n()) %>% 
                                    arrange(desc(Count)))
knitr::kable(Maximum_complaints, "pipe", align=c("l", "c"))
```

### After sorting, we find out that Georgia has the most complaints with a total of 288.




# Which state has the highest percentage of unresolved complaints?
### State with highest percentage of unresolved complaints 
```{r}
Highest_unresolved_Complaints <- sample_data %>% 
  group_by(State) %>% 
  filter(New_Status == "Open") %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))
knitr::kable(Highest_unresolved_Complaints, "pipe", align=c("l", "c","c"))
```

### The state with the highest percentage of unresolved complaints is Georgia with 15.47%.

### State with highest percentage of resolved complaints
```{r}
Highest_resolved_Complaints <- sample_data %>% 
  group_by(State) %>% 
  filter(New_Status == "Closed") %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage =formattable::percent (Count/sum(Count))) %>% 
  arrange(desc(Percentage))
knitr::kable(Highest_resolved_Complaints, "pipe", align=c("l", "c","c"))
```

### The state with the highest percentage of resolved complaints is Georgia with 12.19%.



# Provide the percentage of complaints resolved till date, which were received through the Internet and customer care calls.
### Resolved complaints till date based on based on how they were received
```{r}
Resolved_complaints_till_date <- sample_data%>% 
  group_by(Received.Via) %>% 
  filter(New_Status == "Closed" ) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))
```

### Chart showing total complaints resolved till date  based on how they were received
```{r}
(g4 <- ggplot(data =Resolved_complaints_till_date, aes(x="",y=Percentage, fill = Received.Via ) ) +
    geom_bar(width = 1, stat = "identity") + theme_classic() + coord_polar("y", start=0) + theme(axis.line = element_blank()) + theme(axis.text = element_blank()) + theme(axis.ticks = element_blank()) +
    labs(x = NULL, y = NULL , fill = "Received.Via", title = "Figure 4 Total Complaints Resolved Till Date"   ) + geom_text(aes(label = paste0(Percentage)), position= position_stack(vjust = 0.5)))
```

### It is clear from the pie chart that the total percentage of resolved complaints received via Customer care is 50.62% and those  received via Internet is 49.38%.





### Unresolved Complaints till date  based on how they were received
```{r}
Unresolved_complaints_till_date <- sample_data%>% 
  group_by(Received.Via) %>% 
  filter(New_Status == "Open" ) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))
```

### Graph showing Unresolved Complaints till date  based on how they were received
```{r}
(g5 <- ggplot(data =Unresolved_complaints_till_date, aes(x="",y=Percentage, fill = Received.Via ) ) +
    geom_bar(width = 1, stat = "identity") + theme_classic() + coord_polar("y", start=0) + theme(axis.line = element_blank()) + theme(axis.text = element_blank()) + theme(axis.ticks = element_blank()) +
    labs(x = NULL, y = NULL , fill = "Received.Via", title = "Figure 5 Total Complaints Unresolved Till Date"   ) + geom_text(aes(label = paste0(Percentage)), position= position_stack(vjust = 0.5)))
```

### It is clear from the pie chart that the total percentage of unresolved complaints received via Customer care is 49.32% and those  received via Internet is 50.68%.





### Total status percentage of complaints per how they were received (Internet) 
```{r}
internet_only <- sample_data%>% 
  group_by(New_Status) %>% 
  filter(Received.Via == "Internet" ) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))
```

### Graph showing status percentage of total complaints per how they were received (Internet) 
```{r}
(g6 <- ggplot(data =internet_only, aes(x="",y=Percentage, fill = New_Status ) ) +
    geom_bar(width = 1, stat = "identity") + theme_classic() + coord_polar("y", start=0) + theme(axis.line = element_blank()) + theme(axis.text = element_blank()) + theme(axis.ticks = element_blank()) +
    labs(x = NULL, y = NULL , fill = "Received.Via", title = "Figure 6 Total  complaints per Received Via (Internet) "   ) + geom_text(aes(label = paste0(Percentage)), position= position_stack(vjust = 0.5)))
```


### Total  status percentage of complaints per how they were received (Customer Care)
```{r}
Customer_Care <- sample_data%>% 
  group_by(New_Status) %>% 
  filter(Received.Via == "Customer Care Call" ) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))
```

### Graph showing Total status percentage of complaints per how they were received (Customer Care)
```{r}
(g7 <- ggplot(data =Customer_Care, aes(x="",y=Percentage, fill = New_Status ) ) +
    geom_bar(width = 1, stat = "identity") + theme_classic() + coord_polar("y", start=0) + theme(axis.line = element_blank()) + theme(axis.text = element_blank()) + theme(axis.ticks = element_blank()) +
    labs(x = NULL, y = NULL , fill = "Received.Via", title = "Figure 7 Total  complaints per Received Via (Customer Care) "   ) + geom_text(aes(label = paste0(Percentage)), position= position_stack(vjust = 0.5)))
```



# INSIGHTS
### -The top 3 days with the higest number of complaints in a month are 6th, 24th and 23rd.
### -The Sum of complaints are highest in the month of June.
### -From the table with the frequency of complaint types above, we can see that the maximum complaint type is Others (779), followed by Internet, Billing, Service, and Email.
### -Georgia has the most complaints with a total of 288.
### -The state with the highest percentage of unresolved complaints is Georgia with 15.47%.
### -The state with the highest percentage of resolved complaints is Georgia with 12.19%.
### -The total percentage of resolved complaints received via Customer care is 50.62% and those received via Internet is 49.38%.
### -The total percentage of unresolved complaints received via Customer care is 49.32% and those received via Internet is 50.68%.




